Toward single particle reconstruction without particle picking

### 2018-05-21 TEX/draft4.tex
New try (and ignoring auxiliary files)

@@

\newcommand{\reals}{\mathbb{R}}
\newcommand{\RL}{\mathbb{R}^L}
\newcommand{\CL}{\mathbb{C}^L}
\newcommand{\RN}{\mathbb{R}^N}
\newcommand{\RNN}{\mathbb{R}^{N\times N}}
\newcommand{\CNN}{\mathbb{C}^{N\times N}}
\newcommand{\inner}[1]{\left\langle {#1} \right\rangle}
\newcommand{\hx}{\hat{x}} 
\newcommand{\one}{\mathbf{1}} 
\newcommand{\SNR}{{\textsf{SNR}}}

@@

Single particle reconstruction (SPR) using cryo--electron microscopy (cryo--EM) is an innovative technology for reconstructing the 3-D structure of macromolecules. In recent years, structures
of many molecules, previously regarded as insurmountable, are now being reconstructed to near-atomic resolution; see for instance~\cite{kuhlbrandt2014resolution,bartesaghi20152,smith2014beyond}. This technological advancement was recognized by the 2017 Nobel Prize in Chemistry~\cite{nobel}.

In a cryo--EM experiment, multiple biological samples of the (ideally) same molecule are rapidly frozen in a thin layer of vitreous ice. Within the ice, the molecules are randomly oriented and positioned. Then, the microscope produces a 2D tomographic projection image, called a \emph{micrograph}, of the multiple samples embedded in the ice. The micrograph is dominated by noise due to the small electron doses that
can be applied to the specimen without causing radiation damage.
The cryo--EM problem is to estimate the structure of the molecule from the micrograph (or, typically, several micrographs).

\paragraph{Particle picking.}
All contemporary methods in the field split the reconstruction procedure to several subroutines. 
The first step in the pipeline is the so-called \emph{particle picking}, in which one aims to detect the 2-D tomographic projections of the samples from the noisy micrograph. The output of ideal particle picker is a series of 2-D images, each containing one centered tomographic projection associated with an unknown viewing direction and an in-plane rotation. This series of images 
is then used to estimate the structure.

The significance of the model bias in cryo--EM was stressed in~\cite{shatsky2009method,henderson2013avoiding} and demonstrated by the ``Einstein from noise'' example.
In this experiment, the image of Einstein is correlated with multiple images of pure noise. The noisy images are aligned to Einstein's image using cross-correlation and then averaged. 
In Figure~\ref{fig:EinsteinFromNoise} we conducted this experiment with different number of noise images. With merely 1000 noise images, Einstein's face is clearly observed in the averaged image. With $10,000$  image, the output is remarkably similar to the template image of Einstein rather than to pure noise. Crucially, the more noisy images we use, the more severe the model bias towards the template. 
In the context of cryo--EM, this example demonstrates how our prior assumptions about the particle may influence the reconstructed structure.
This true in general to any method which is based on template matching.

\begin{figure}[h!]
	\centering
	\begin{subfigure}[h]{0.3\textwidth}
		\centering
		\includegraphics[scale=0.3]{Einstein}
		\caption{\label{fig:Einstein}Einstein image (template)}
	\end{subfigure}%
	\begin{subfigure}[h]{0.3\textwidth}
		\centering
		\includegraphics[scale=0.3]{Einstein_from_noise_n10}
		\caption{$n = 10$}
	\end{subfigure}
	\begin{subfigure}[h]{0.3\textwidth}
		\centering
		\includegraphics[scale=0.3]{Einstein_from_noise_n100}
		\caption{$n = 10^2$}
	\end{subfigure}

	\begin{subfigure}[h]{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{Einstein_from_noise_n1000}
	\caption{$n = 10^3$}
\end{subfigure}%
\begin{subfigure}[h]{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{Einstein_from_noise_n10000}
	\caption{$n = 10^4$}
\end{subfigure}
\begin{subfigure}[h]{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{Einstein_from_noise_n100000}
	\caption{$n = 10^5$}
\end{subfigure}

	\caption{\label{fig:EinsteinFromNoise} Einstein from noise experiment. As a template, we used Einstein's image of size $50\times 50$ pixels. We drawn $n$ images of the same size whose entries are composed of i.i.d.\ Gaussian entries with the same mean and variance as Einstein's image. Each noise image was aligned with respect to the template by the maximal value of their cross correlation. After the alignment, all the images were averaged. As can be seen, the more pure noise images we use, the closer the averaged image to the template.   }	
\end{figure}

@@

Figure~\ref{fig:micro_example} shows a simple example of our problem with  one signal ($K=1$). 
Each panel demonstrates an image of size $250\times 250$, containing 4 occurrences of an Einstein image of size $50\times 50$ presented in Figure~\ref{fig:Einstein}. 
The left panel shows the data without noise. In the middle panel,  i.i.d.\ Gaussian noise with standard deviation $\sigma=0.5$  was added, however,  Einstein is still easily identified. In this scenario it is likely that standard detection algorithms can locate the copies of the signal within the micrograph. In the right panel, we show the same data swamped in  
i.i.d.\ Gaussian noise with standard deviation  $\sigma=3$.
Clearly, identifying the locations of Einstein in this micrograph is much more challenging. In fact, it can be shown that in the low signal--to--noise (\SNR) regime, 
detection  of individual image occurrences is impossible---even if the true image is known---and therefore particle picking is impossible.  This phenomenon is explained in detail in Appendix~\ref{sec:theory}.

\begin{figure}[h!]
	\centering
	\begin{subfigure}[h]{0.33\textwidth}
		\centering
		\includegraphics[scale=0.5]{micrograph_Einstein_example_clean}
		\caption{$\sigma = 0$}
	\end{subfigure}%
	\begin{subfigure}[h]{0.33\textwidth}
		\centering
		\includegraphics[scale=0.5]{micrograph_Einstein_example_s05}
		\caption{$\sigma = 0.5$}
	\end{subfigure}
	\begin{subfigure}[h]{0.33\textwidth}
		\centering
		\includegraphics[scale=0.5]{micrograph_Einstein_example_s3}
		\caption{$\sigma = 3$}
	\end{subfigure}
	\caption{\label{fig:micro_example} Example of data sequences (micrographs) of size $250\times 250$ with additive i.i.d.\ Gaussian noise with variance $\sigma^2$. Each micrograph contains four occurrences of a $50 \times 50$ image of Einstein; see Figure~\ref{fig:Einstein}.}	
\end{figure}

@@

\paragraph{Mathematical model}

@@

\paragraph{Estimating the autocorrelation function of a single signal}

@@

\paragraph{Estimating the autocorrelation function of a multiple signals}

@@

\paragraph{Numerical experiments details}
For the 2-D experiment shown in Figures~\ref{fig:Einst_example} and~\ref{fig:error_per_micro}, we generated $P$ micrographs of size $4096\times 4096$ pixels. 
In each micrograph, we placed Einstein's image of size $50\times 50$  in random locations (see Figure~\ref{fig:Einstein}), while keeping the separation condition~\eqref{eq:spacing}.  
This is done by randomly selecting $4000$ placements in the micrograph, one at a time with
an accept/reject rule based on the separation constraint and locations picked so far. In average, $700$ images were placed in each micrograph.   
Then, an i.i.d.\ Gaussian noise with standard deviation $\sigma=3$ was added inducing $\SNR$ of approximately $1/20$. An example of a micrograph's excerpt is presented in the right panel of Figure~\ref{fig:micro_example}.
%Different micrographs are handled sequentially on a GPU, as GPUs are particularly well suited to execute simple instructions across large vectors of data. If multiple GPUs are available, segments can of course be handled in parallel.

@@

\section{Autocorrelation estimations} \label{sec:autocorrelation_computation}

Throughout the proof, we consider the case of one signal $K=1$. The extension to $K>1$ is straightforward by averaging the contributions of all signal with  appropriates weights, see~\cite{boumal2017heterogeneous}.

Let us define
\begin{equation}
\gamma = \lim_{N\to\infty} \frac{M_NL}{N}<1.
\end{equation}
With a bit abuse of notation, $M_N$ stresses that $M$ is a function of $N$. Indeed, by assuming $M_N=\Omega(N)$, we deduce $\gamma>0$.
We start by considering the first autocorrelation of the data
\begin{equation}
a_y^1 = \sum_{i=0}^{N-1} y[i] = \frac{1}{N/L}\sum_{j=0}^{M_N-1}\frac{1}{L}\sum_{i=0}^{L-1}x[i] + \underbrace{\frac{1}{N}\sum_{i=0}^{N-1}\varepsilon[i]}_{\text{noise term}} \xrightarrow{a.s.}\gamma a_x^1,
\end{equation}
where the noise term converges to zero almost surely (a.s.) by the law of large numbers.

We proceed with the second autocorrelation for fixed $\ell\in[0,\ldots,L-1]$. We can compute:
\begin{equation}
\begin{split}
a_y^2[\ell] & = \frac{1}{N}\sum_{i=0}^{N-1-\ell}y[i]y[i+\ell] \\
& \underbrace{\frac{1}{N}\sum_{j=1}^{M_N}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]}_{\text{signal term}} + \underbrace{\frac{1}{N}\sum_{i=0}^{N-1-\ell}\varepsilon[i]\varepsilon[i+\ell]}_{\text{noise term}},
\end{split}
\end{equation}
where the cross terms between the signal and the noise vanish  almost surely in the limit $N\to\infty$.

We treat the signal and noise terms separately. We first break the signal term into $M_N$ different sums, each contains one copy of the signal, and get
\begin{equation} \label{eq:2nd_moment_signal_term}
\frac{1}{N}\sum_{j=1}^{M_N}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell] = \frac{M_NL}{N}\frac{1}{L}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell] = \gamma a_x^2[\ell].
\end{equation}
Similarly, for $\ell\neq 0$, we can break the noise term into a sum of independent terms 
\begin{equation}
\frac{1}{N}\sum_{i=0}^{N-1-\ell} \varepsilon[i]\varepsilon[i+\ell] = \frac{1}{\ell}\sum_{i=0}^{\ell-1}\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -1} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i].
\end{equation}
Each term of $\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -1} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i]$ is an average of $N/\ell$ independent terms with expectation zero, and thus converges to zero almost surely as $N\to\infty$.
If $\ell=0$, 
\begin{equation}
\frac{1}{N}\sum_{i=0}^{N-1} \varepsilon^2[i] \xrightarrow{a.s.} \sigma^2.
\end{equation}

We are now moving to analyze the third-order autocorrelation. Let us fix $\ell_1\geq\ell_2$ and recall that 
\begin{equation*}
a_y^3[\ell_1,\ell_2] = \sum_{i=0}^{N-1-\ell_1} y[i]y[i+\ell_1]y[i+\ell_2]. 
\end{equation*}
Writing explicitly in terms of signal and noise, the sum can be broken into eight terms. The first contains only signal terms (does not see noise) and converges to $\gamma a_x^3$ from the same reasons as~\eqref{eq:2nd_moment_signal_term}. Three other terms contain the product of two signal entries and one noise term. Since the noise is independent of the signal and has zero mean, these terms go to zero almost surely.

@@

\section{Related literature} \label{sec:related_literature}

\paragraph{Blind deconvolution.}
Blind deconvolution is a longstanding problem, arising in a variety of engineering and scientific applications, such as astronomy, communication, image deblurring, system identification and optics; see~\cite{jefferies1993restoration,shalvi1990new,ayers1988iterative,abed1997blind}, just to name a few. 
To make the problem well-posed, we must  assume some prior knowledge or  structure. 
In our case, the prior information is that $s$ is a binary signal that satisfies the separation constraint~\eqref{eq:spacing}. 
Other settings of blind deconvolution problems have been analyzed recently, see for instance~\cite{ahmed2014blind,li2016identifiability,li2016rapid,ling2015self,ling2017blind,chi2016guaranteed}
where the focus is on high $\SNR$ regimes.

An important feature of the problem under consideration is that while both $x_i$'s and $s_i$'s are unknown, the goal is merely to estimate the $x_i$'s. The  $s_i$'s  are referred to as \emph{nuisance  variables}. Indeed, in many blind deconvolution applications the sole purpose is to recover one of the unknown signals. For instance, in image deblurring, both the blurring
kernel and the high-resolution image are unknown, but the primary goal is only
to sharpen the image.

\paragraph{The impossibility of detection in low $\SNR$.}
If $x$ is known and $K=1$, then a sparse signal can be estimated by linear programming  in the high $\SNR$ regime, e.g.,~\cite{azais2015spike,denoyelle2017support,bendory2016robust,bendory2017robust,bernstein2017deconvolution}. However, in the low $\SNR$ regime, estimating the binary sparse signal $s$ is impossible. To see that, suppose that an oracle provides us $M$ windows of length $W>L$, each contains one copy of $x$. That is to say, we get a series of windows of length $W$, each one contains a signal at an unknown location. 
Estimating the position of the known signal within each  window is an easier problem than detecting the support of $s$. 
Nevertheless, even this problem is impossible in the low $\SNR$ regime~\cite{aguerrebere2016fundamental}. Therefore, we conclude that detecting the nonzero values of $s$ is impossible in low $\SNR$. As aforementioned, this work focuses on this regime and examines under what conditions we can estimate the signals, despite the impossibility of detecting their individual occurrences.

\paragraph{System identification.}
For $K=1$, our problem can be also interpreted as a special case of the system identification problem. Similarly to~\eqref{eq:model}, the system identification forward model takes the form
\begin{eqnarray}
y = x\ast w + \varepsilon,  
\end{eqnarray} 
where $x$ is the unknown signal (``system''), $w$ is an unknown, random, input sequence and $\varepsilon$ is an additive noise.   
%The problem has also been studied in the case of a known input $w$~\cite{pillonetto2010new,dinuzzo2015kernels,bottegal2016robust}. 
The goal of this problem is to estimate $x$, usually referred to as ``identifying the system.'' The question of identifiability of $x$ under
this observation model is addressed for certain Gaussian and non-Gaussian $w$ in~\cite{benveniste1980robust,kormylo1983identifiability}.
In the special case where $w\in\{0,1\}^N$, satisfying the spacing requirement~\eqref{eq:spacing}, we obtain our
model in the  case of a single signal ($K = 1$).

>>>>

\newcommand{\LL}{\mathcal{L}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\ep}{\varepsilon}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\GCD}{\mathbf{GCD}}
\newcommand{\XX}{\mathcal{X}}
\newcommand{\SUM}{\text{sum}}
\newcommand{\1}{\mathbf{1}}

@@

\newcommand{\reals}{\mathbb{R}}
\newcommand{\RL}{\mathbb{R}^L}
\newcommand{\CL}{\mathbb{C}^L}
\newcommand{\RN}{\mathbb{R}^N}
\newcommand{\RNN}{\mathbb{R}^{N\times N}}
\newcommand{\CNN}{\mathbb{C}^{N\times N}}
\newcommand{\inner}[1]{\left\langle {#1} \right\rangle}
\newcommand{\hx}{\hat{x}} 
\newcommand{\one}{\mathbf{1}} 
\newcommand{\SNR}{\ensuremath{\textsf{SNR}}}

@@

Cryo--electron microscopy (cryo--EM) is an innovative technology for single particle reconstruction (SPR) of macromolecules. 
%In recent years, structures of many molecules, previously regarded as insurmountable, are now being reconstructed to near-atomic resolution
%; see for instance~\cite{kuhlbrandt2014resolution,bartesaghi20152,smith2014beyond}.
%This technological advancement was recognized by the 2017 Nobel Prize in Chemistry~\cite{nobel}. 
% 
In a cryo--EM experiment, biological samples are rapidly frozen in a thin layer of vitreous ice. Within the ice, the molecules are randomly oriented and positioned. The microscope produces a 2D image of the samples embedded in the ice called a \emph{micrograph}. Each micrograph contains tomographic projections of the samples at unknown locations and under unknown viewing directions.

The signal to noise ratio ($\SNR$) of the projections in the micrographs is a function of two dominating factors. On the one hand, the $\SNR$ is a function of the electron dose. To keep radiation damage within acceptable bounds, the dose must be kept low, which leads to high noise levels. On the other hand, the $\SNR$ is a function of the molecule size. The smaller the molecule, the fewer detected electrons carry information about the sample.

%\paragraph{Particle picking.}
All contemporary methods in the field split the reconstruction procedure in several stages.
The first stage consists in extracting the various particle projections from the micrographs. This is called \emph{particle picking}. Ulterior stages aim to construct a 3-D model of the molecule from these projections. The quality of the reconstruction eventually hinges on the quality of the particle picking stage. As can be seen from Figure~\ref{fig:micro_example}, locating the particles becomes increasingly challenging as the $\SNR$ degrades. Crucially, it can be shown that detection of individual particles is impossible below a certain critical $\SNR$---\TODO{ref}.

This fact has been recognized early on by the cryo-EM community. In particular, in a highly cited paper from 1995, Henderson~\cite{henderson1995limitations} investigates the following questions:
\begin{quote}
	\emph{For the purposes of this review, I would like to ask the question: what is the smallest size of free-standing molecule whose structure can in principle be determined by phase-contrast electron microscopy? Given what has already been demonstrated in published work, this reduces to the question: what is the smallest size of molecule for which it is possible to determine from images of unstained molecules the five parameters needed to define accurately its orientation (three parameters) and position (two parameters) so that averaging can be performed?}
\end{quote}
In this paper and in others that followed~[gaebner?], it was established that particle picking is impossible for molecules below a certain weight. This has been widely understood as meaning that SPR itself is impossible for such molecules, as exemplified by recent literature~[lots of things here.]

In this paper, we argue that there is a gap between the two questions in the quoted excerpt above. Specifically, the impossibility of particle picking does not necessarily imply impossibility of particle reconstruction. This can be understood from a statistical point of view by counting the number of parameters that one aims to estimate. For a given target resolution, the number of parameters required to describe the 3-D model of the molecule is fixed: it does not depend on the amount of data collected. In contrast, particle picking involves estimation of several so-called nuisance parameters for each projection present in the micrographs, such as their locations for example. Thus, acquiring more data does not necessarily allow to improve the particle picking stage, whereas it should in principle allow to improve the molecule reconstruction.

The significance of the model bias in cryo--EM was stressed in~\cite{shatsky2009method,henderson2013avoiding} and demonstrated by the ``Einstein from noise'' example.
In this experiment, the image of Einstein is correlated with multiple images of pure noise. The noisy images are aligned to Einstein's image using cross-correlation and then averaged. 
In~\cite{shatsky2009method}, it was shown that with merely  1000 noise images, the averaged image is remarkably similar to the template image of Einstein rather than to pure noise.
%In Figure~\ref{fig:EinsteinFromNoise} we conducted this experiment with different number of noise images. With merely 1000 noise images, Einstein's face is clearly observed in the averaged image. With $10,000$  image, the output is remarkably similar to the template image of Einstein rather than to pure noise. Crucially, the more noisy images we use, the more severe the model bias towards the template. 
In the context of cryo--EM, this example demonstrates how our prior assumptions about the particle may influence the reconstructed structure.
This true in general to any method which is based on template matching.
\TODO{Do we have a nice experiment to exemplify the phenomenon?}

%\begin{figure}[h!]
%	\centering
%	\begin{subfigure}[h]{0.3\textwidth}
%		\centering
%		\includegraphics[scale=0.3]{Einstein}
%		\caption{\label{fig:Einstein}Einstein image (template)}
%	\end{subfigure}%
%	\begin{subfigure}[h]{0.3\textwidth}
%		\centering
%		\includegraphics[scale=0.3]{Einstein_from_noise_n10}
%		\caption{$n = 10$}
%	\end{subfigure}
%	\begin{subfigure}[h]{0.3\textwidth}
%		\centering
%		\includegraphics[scale=0.3]{Einstein_from_noise_n100}
%		\caption{$n = 10^2$}
%	\end{subfigure}

%	\begin{subfigure}[h]{0.3\textwidth}
%	\centering
%	\includegraphics[scale=0.3]{Einstein_from_noise_n1000}
%	\caption{$n = 10^3$}
%\end{subfigure}%
%\begin{subfigure}[h]{0.3\textwidth}
%	\centering
%	\includegraphics[scale=0.3]{Einstein_from_noise_n10000}
%	\caption{$n = 10^4$}
%\end{subfigure}
%\begin{subfigure}[h]{0.3\textwidth}
%	\centering
%	\includegraphics[scale=0.3]{Einstein_from_noise_n100000}
%	\caption{$n = 10^5$}
%\end{subfigure}

%\caption{\label{fig:EinsteinFromNoise} Einstein from noise experiment. As a template, we used Einstein's image of size $50\times 50$ pixels. We drawn $n$ images of the same size whose entries are composed of i.i.d.\ Gaussian entries with the same mean and variance as Einstein's image. Each noise image was aligned with respect to the template by the maximal value of their cross correlation. After the alignment, all the images were averaged. As can be seen, the more pure noise images we use, the closer the averaged image to the template.   }	
%\end{figure}

@@

Figure~\ref{fig:micro_example} shows a simple example of our problem with  one signal ($K=1$). 
Each panel demonstrates an image of size $250\times 250$, containing 4 occurrences of an Einstein image of size $50\times 50$. 
The left panel shows the data without noise. In the middle panel,  i.i.d.\ Gaussian noise with standard deviation $\sigma=0.5$  was added, however,  Einstein is still easily identified. In this scenario it is likely that standard detection algorithms can locate the copies of the signal within the micrograph. In the right panel, we show the same data swamped in  
i.i.d.\ Gaussian noise with standard deviation  $\sigma=3$.
Clearly, identifying the locations of Einstein in this micrograph is much more challenging. In fact, it can be shown that in the low signal--to--noise ($\SNR$) regime, 
detection  of individual image occurrences is impossible---even if the true image is known---and therefore particle picking is impossible.  This phenomenon is explained in detail in Appendix~\ref{sec:theory}.

\begin{figure}[h!]
	\centering
	\begin{subfigure}[h]{0.33\textwidth}
		\centering
		\includegraphics[scale=0.5]{micrograph_Einstein_example_clean}
		\caption{$\sigma = 0$}
	\end{subfigure}%
	\begin{subfigure}[h]{0.33\textwidth}
		\centering
		\includegraphics[scale=0.5]{micrograph_Einstein_example_s05}
		\caption{$\sigma = 0.5$}
	\end{subfigure}
	\begin{subfigure}[h]{0.33\textwidth}
		\centering
		\includegraphics[scale=0.5]{micrograph_Einstein_example_s3}
		\caption{$\sigma = 3$}
	\end{subfigure}
	\caption{\label{fig:micro_example} Example of data sequences (micrographs) of size $250\times 250$ with additive i.i.d.\ Gaussian noise with variance $\sigma^2$. Each micrograph contains four occurrences of a $50 \times 50$ image of Einstein.}	
\end{figure}

@@

\paragraph{Mathematical model.}

@@

\paragraph{Estimating the autocorrelation function of a single signal.}

@@

\paragraph{Estimating the autocorrelation function of a multiple signals.}

@@

\paragraph{Numerical experiments details.}
For the 2-D experiment shown in Figures~\ref{fig:Einst_example} and~\ref{fig:error_per_micro}, we generated $P$ micrographs of size $4096\times 4096$ pixels. 
In each micrograph, we placed Einstein's image of size $50\times 50$  in random locations, while keeping the separation condition~\eqref{eq:spacing}.  
This is done by randomly selecting $4000$ placements in the micrograph, one at a time with
an accept/reject rule based on the separation constraint and locations picked so far. In average, $700$ images were placed in each micrograph.   
Then, an i.i.d.\ Gaussian noise with standard deviation $\sigma=3$ was added inducing $\SNR$ of approximately $1/20$. An example of a micrograph's excerpt is presented in the right panel of Figure~\ref{fig:micro_example}.
%Different micrographs are handled sequentially on a GPU, as GPUs are particularly well suited to execute simple instructions across large vectors of data. If multiple GPUs are available, segments can of course be handled in parallel.

@@

\section{Proof of \eqref{eq:data_ac}} \label{sec:autocorrelation_computation}

Throughout the proof, we consider the case of one signal $K=1$. The extension to $K>1$ is straightforward by averaging the contributions of all signal with  appropriate weights; see~\cite{boumal2017heterogeneous}.

We will let the number of instances of the signal $M$ grow with $N$, and write $M=M_N$ to emphasize this. We assume $M_N$ grows proportionally with $N$, and define:
%
\begin{equation}
\gamma = \lim_{N\to\infty} \frac{M_NL}{N}<1.
\end{equation}
%
We will assume that $M_N=\Omega(N)$, so that $\gamma>0$. In the sequel, we will suppress the explicit dependence of $M$ on $N$ for notational convenience.

We start by considering the mean of the data:
%
\begin{equation}
a_y^1 = \frac{1}{N}\sum_{i=0}^{N-1} y[i] =
\frac{1}{N/L}\sum_{j=0}^{M-1}\frac{1}{L}\sum_{i=0}^{L-1}x[i] +    
\underbrace{\frac{1}{N}\sum_{i=0}^{N-1}\varepsilon[i]}_{\text{noise term}}
\xrightarrow{a.s.}\gamma a_x^1,
\end{equation}
%
where the noise term converges to zero almost surely (a.s.) by the strong law of large numbers.

We proceed with the (second-order) autocorrelation for fixed $\ell\in[0,\ldots,L-1]$. We can compute:
%

\begin{align}
%
a_y^2[\ell] & = \frac{1}{N}\sum_{i=0}^{N-1-\ell}y[i]y[i+\ell]
\nonumber \\
& = \underbrace{\frac{1}{N}\sum_{j=1}^{M}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]}_{\text{signal term}} + \underbrace{\frac{1}{N}\sum_{i=0}^{N-1-\ell}\varepsilon[i]\varepsilon[i+\ell]}_{\text{noise term}}
+ \underbrace{\frac{1}{N} \sum_{j=1}^{M} \sum_{i=0}^{L-1} x[i] \varepsilon[s_j + i + \ell]}_{\text{cross-term}}.
%
\end{align}

The cross-terms are linear in the noise, and are easily shown to vanish almost surely in the limit $N\to\infty$, by the strong law of large numbers. As for the signal term, we break it into $M$ different sums, each containing one copy of the signal. This gives:
%
\begin{equation} \label{eq:2nd_moment_signal_term}
%
\frac{1}{N}\sum_{j=1}^{M}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell] = \frac{ML}{N}\frac{1}{L}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]\xrightarrow{N\to\infty}\gamma a_x^2[\ell].
%
\end{equation}
%

We next analyze the pure noise term. When $\ell\neq 0$, we can break the noise term into a sum of independent terms:
%
\begin{equation}
%
\frac{1}{N}\sum_{i=0}^{N-1-\ell} \varepsilon[i]\varepsilon[i+\ell] = \frac{1}{\ell}\sum_{i=0}^{\ell-1}\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -1} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i].
%
\end{equation}
%
Each sum $\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -1} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i]$ is an average of $N/\ell$ independent terms with expectation zero, and thus converges to zero almost surely as $N\to\infty$. If $\ell=0$, then we have:
%
\begin{equation}
%
\frac{1}{N}\sum_{i=0}^{N-1} \varepsilon^2[i] \xrightarrow{a.s.} \sigma^2.
%
\end{equation}

We now analyze the third-order autocorrelation. Let us fix $\ell_1\geq\ell_2\ge0$. We have:
%
\begin{align}
%
&a_y^3[\ell_1,\ell_2] 
= \frac{1}{N}\sum_{i=0}^{N-1-\ell_1} y[i]y[i+\ell_1]y[i+\ell_2]
\nonumber \\
%
=& \underbrace{ \frac{ML}{N}\frac{1}{M}\sum_{j=1}^M 
	\frac{1}{L}\sum_{i=0}^{L-1-\ell_1}x[i]x[i+\ell_1]x[i+\ell_2]   }_{(1)}
+ \underbrace{\frac{1}{N}\sum_{i=0}^{N-1-\ell_1} \ep[i]\ep[i+\ell_1]\ep[i+\ell_2]}_{(2)}
\nonumber \\
&+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-1} x[i]\ep[s_j + i+\ell_1]\ep[s_j+ i+\ell_2]}_{(3)}
+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-1} \ep[s_j+i-\ell_1]x[i]\ep[s_j+ i+\ell_2-\ell_1]}_{(4)}
\nonumber \\
&+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-1} \ep[s_j+i-\ell_2]\ep[s_j+i+\ell_1-\ell_2]x[i]}_{(5)}
+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-\ell_1+\ell_2-1} \ep[s_j+i]x[i+\ell_1-\ell_2]x[i]}_{(6)}
\nonumber \\
&+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-\ell_2-1} x[i]\ep[s_j + i+\ell_1]x[s_j+ i+\ell_2]}_{(7)}
+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-\ell_1-1} x[i]x[i+\ell_1]\ep[s_j+ i+\ell_2]}_{(8)}.
%
\end{align}
%
Terms (6), (7) and (8) are linear in $\ep$, and can easily be shown to converge to 0 almost surely by the law of large numbers, by similar arguments as used previously. Term (1) converges to $\gamma a_x^3[\ell_1,\ell_2]$ almost surely, for the same reasons as~\eqref{eq:2nd_moment_signal_term}. To deal with terms (2)--(5), we must distinguish between different values of $\ell_1$ and $\ell_1$.

{\bf Case 1:} $0 < \ell_2 < \ell_1$. Here, all summands with elements of $\ep$ involve products of distinct entries, which have expected value 0. Consequently, the usual argument shows that terms (2)--(5) all converge to 0 almost surely as $N \to \infty$.

{\bf Case 2:} $0=\ell_2 < \ell_1$. Term (2) is an average of products of the form $\ep[i]^2\ep[i+\ell_1]$, which have mean zero; consequently, term (2) converges to 0 almost surely. The same argument as for Case 1 shows that (3) and (5) also converge to 0. For term (4), we write:
%
\begin{align}
%
&\frac{1}{N}\sum_{j=1}^{M} 
\sum_{i=0}^{L-1} \ep[s_j+i-\ell_1]x[i]\ep[s_j+ i+\ell_2-\ell_1]
\nonumber \\
&= \frac{ML}{N}\frac{1}{L}\sum_{i=0}^{L-1}x[i] \frac{1}{M}\sum_{j=1}^{M} \ep[s_j+i-\ell_1]^2
\xrightarrow{N\to\infty} \gamma \frac{1}{L} \sum_{i=0}^{L-1}x[i] \sigma^2 = \gamma a_x^1 \sigma^2.
%
\end{align}

{\bf Case 3:} $0<\ell_2 = \ell_1$. An argument nearly identical to that for Case 2 shows that terms (2), (4) and (5) converge to 0, while term (3) converges to $\gamma a_x^1 \sigma^2$.

{\bf Case 4:} $0=\ell_2 = \ell_1$. The same argument as for term (4) in Case 2 shows that terms (3), (4) and (5) all converge to $\gamma a_x^1 \sigma^2$. Term (2) is an average of $\ep[i]^3$, which is mean zero; consequently, it converges to 0.

This completes the proof of \eqref{eq:data_ac}.

\begin{comment}
Writing explicitly in terms of signal and noise, the sum can be broken into eight terms. The first contains only signal terms (does not see noise) and converges to $\gamma a_x^3$ from the same reasons as~\eqref{eq:2nd_moment_signal_term}. Three other terms contain the product of two signal entries and one noise term. Since the noise is independent of the signal and has zero mean, these terms go to zero almost surely.

@@

\end{comment}

\section{Related work} \label{sec:related_literature}

\paragraph{System identification.}

For $K=1$, our problem can be interpreted as a special case of the system identification problem. Similarly to~\eqref{eq:model}, the system identification forward model takes the form
%
\begin{math}
%
y = x\ast w + \varepsilon,  
%
\end{math} 
%
where $x$ is the unknown signal (the ``system''), $w$ is an unknown, random, input sequence, and $\varepsilon$ is an additive noise.   
%The problem has also been studied in the case of a known input $w$~\cite{pillonetto2010new,dinuzzo2015kernels,bottegal2016robust}. 
The goal of this problem is to estimate $x$, usually referred to as ``identifying the system.'' The question of identifiability of $x$ under this observation model is addressed for certain Gaussian and non-Gaussian $w$ in~\cite{benveniste1980robust,kormylo1983identifiability}. In the special case where $w\in\{0,1\}^N$, satisfying the spacing requirement~\eqref{eq:spacing}, we obtain our model in the  case of a single signal ($K = 1$). The same observation model is used for blind deconvolution, a longstanding problem arising in a variety of engineering and scientific applications such as astronomy, communication, image deblurring, system identification and optics; see~\cite{jefferies1993restoration,shalvi1990new,ayers1988iterative,abed1997blind}, just to name a few.

%%%To make the problem well-posed, we must  assume some prior knowledge or  structure.  In our case, the prior information is that $s$ is a binary signal that satisfies the separation constraint~\eqref{eq:spacing}.  Other settings of blind deconvolution problems have been analyzed recently, see for instance~\cite{ahmed2014blind,li2016identifiability,li2016rapid,ling2015self,ling2017blind,chi 016guaranteed} where the focus is on high $\SNR$ regimes. Importantly, in the system identification problem, the goal is only to recover $x$, not the sequence $w$, or in our case, the signal locations $s_i$. We refer to the $s_i$'s as \emph{nuisance parameters}.

%%%An important feature of the problem under consideration is that while both $x_i$'s and $s_i$'s are unknown, the goal is merely to estimate the $x_i$'s. The  $s_i$'s  are referred to as \emph{nuisance  variables}. Indeed, in many blind deconvolution applications the sole purpose is to recover one of the unknown signals. For instance, in image deblurring, both the blurring kernel and the high-resolution image are unknown, but the primary goal is only to sharpen the image.

\paragraph{Likelihood-based methods.}

Likelihood-based methods estimate $x$ as the maximizer of some function $f(x | y)$, where $f$ is derived from the likelihood function of $x$ given the observed signal $y$. For example, $f$ may be the likelihood itself, or a related function with a similar form (leading to the class of ``quasi-likelihood'' methods). If some prior is assumed on $x$, then $f(x|y)$ can be taken to be the posterior distribution of $x$ given the data; this is the simplest form of Bayesian inference.

Optimizing the function $f(x|y)$ exactly is often intractable, and thus heuristic methods are used instead. One proposed technique is to use Markov Chain Monte Carlo (MCMC)~\cite{cappe1999simulation}. Another paper considers parameterized models for multiple distinct signals, as in our framework ($K>1$)~\cite{andrieu2001bayesian}. Their proposed solution is an MCMC algorithm tailored for their specific parametrized problem.
